---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# graphcoloring

[![Travis build status](https://travis-ci.org/saurfang/graphcoloring.svg?branch=master)](https://travis-ci.org/saurfang/graphcoloring)
[![CRAN status](https://www.r-pkg.org/badges/version/graphcoloring)](https://cran.r-project.org/package=graphcoloring)
[![lifecycle](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![Coverage status](https://codecov.io/gh/saurfang/graphcoloring/branch/master/graph/badge.svg)](https://codecov.io/github/saurfang/graphcoloring?branch=master)

`graphcoloring` is a collection of graph coloring algorithms for coloring vertices of a graph such that no two adjacent vertices share the same color. The algorithms
are included via the embedded 'GraphColoring' C++ library, <https://github.com/brrcrites/GraphColoring>.

## Installation

You can install the released version of graphcoloring from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("graphcoloring")
```

Development version can be installed with

```r
devtools::install_github("saurfang/graphcoloring")
```

## Example

`color_*` functions operate under `tidygraph` family and can be used to color nodes within `mutate` context similar to `group_*` functions in `tidygraph`.

```{r example, message=FALSE}
library(graphcoloring)
library(tidygraph)
library(ggraph)

set.seed(42)

play_islands(5, 10, 0.8, 3) %>%
  mutate(color = as.factor(color_dsatur())) %>%
  ggraph(., layout = 'kk') +
  geom_edge_link(aes(alpha = ..index..), show.legend = FALSE) +
  geom_node_point(aes(color = color), size = 7) +
  theme_graph()
```

`graph_coloring_*` functions directly take adjacency lists and returns an integer vector of assigned labels.
For example, this can be used with `sf::st_intersects()` to color a feature collection for visualization.

```{r sf-example, message=FALSE}
library(graphcoloring)
library(USAboundaries)
library(sf)
library(ggplot2)

set.seed(48)

us_states() %>%  
  filter(!(name %in% c("Alaska", "District of Columbia", "Hawaii", "Puerto Rico"))) %>%
  mutate(
    color = st_intersects(.) %>%
      graph_coloring_dsatur() %>%
      as.factor()
  ) %>%
  ggplot() +
  geom_sf(aes(fill = color)) +
  theme_bw()
```

